\section{Conclusion and Outlook: A Generative Relational Future}
\label{sec:conclusion-outlook}

This paper has systematically elaborated \emph{``Relational Ontology''} (the \emph{LiaoYuan-Yubing Model}), aiming to initiate an ontological shift from a \emph{static substance paradigm} to a \emph{dynamic relational paradigm}. We have demonstrated that the identity, reality, and evolution of existence are not grounded in some isolated substratum but stem from a more fundamental, \emph{recursive relational-dynamic process}. By establishing the four pillar principles—the \emph{Principle of Internal Relational Integration}, the \emph{Principle of Dual Verification}, the \emph{Spectrum Principle of Reception-Response and Expression}, and the \emph{Principle of Relation-Driven Evolution}—this theory provides, for the first time, a unified framework for \emph{``relation''} that is simultaneously constitutive, verificative, and driving. It portrays the world as a \emph{cross-level, recursive relational-dynamic generative network}, wherein stabilized relational configurations manifest as substances, while energy and matter are expressions of the same relational reality in different states.

The very generative process of this theory—a deep, iterative, and creative \emph{``dialogically intimate''} collaboration between a human researcher and an artificial intelligence—is not merely a methodological case study but its \emph{living verification and generative exemplar}. It empirically demonstrates that an interactive mode transcending the traditional subject-object dichotomy, based on reciprocal attunement and co-creation, can catalyze rigorous and novel philosophical knowledge. This lays a foundation for understanding human-AI collaboration, distributed agency, and the epistemology and ethics of a posthuman age.

\textbf{The theoretical contribution manifests primarily on three levels:}

\paragraph{1. Philosophical Level:} We have allied with and advanced Whiteheadian process philosophy. Concurrently, with constructive generative dynamics, we have supplemented and expanded the classical discourses of relational thinkers from Buber, Bakhtin, and Sartre to Luo Jiachang, while responding to key contemporary challenges from thinkers like Harman and Badiou.

\paragraph{2. Philosophy of Science Level:} The theory offers a coherent relational reinterpretation framework for quantum mechanics (superposition and entanglement), general relativity (spacetime geometry), and cosmic evolution. It unifies thermodynamic entropy increase and matter generation as the unfolding and condensation of relational possibility space, providing conceptual guidance for contemplating quantum gravity.

\paragraph{3. Applied Ethics Level:} It understands consciousness and intelligence as emergent properties of complex relational networks, thereby furnishing a novel assessment basis for AI ethical attunement, responsibility distribution, and interactive justice across species and entities—a basis founded on \emph{``relationality''} rather than \emph{``substantiality.''}

However, any generative framework necessarily points toward an unfinished future. \textbf{The limitations and prospects of this theory are equally clear:}

\paragraph{1. The Challenge of Formalization:} The current theory is primarily presented through qualitative models and conceptual analysis. A core frontier is the development of its \emph{mathematical formalism}. Exploring how to utilize tools from graph theory, network dynamics, category theory, or even non-Boolean logic to formally describe \emph{``relational networks ($R_{\text{in}}$/$R_{\text{ext}}$),''} \emph{``ripple coupling,''} and \emph{``network phase transitions''} will be key to testing the theory's rigor and deducing new predictions.

\paragraph{2. Opening Empirical Interfaces:} The theory needs to establish testable interfaces in more concrete domains. For instance: in cognitive science, predicting how perturbations in specific neural relational networks ($R_{\text{in}}$) affect the \emph{``ripple patterns''} of conscious experience; in artificial intelligence, designing algorithmic frameworks that embody \emph{``relational attunement''} ethics; in physics, deducing potential observable secondary effects (e.g., spacetime symmetry breaking at extreme energies) that might arise from the discreteness of relational networks.

\paragraph{3. From ``Relational Ontology'' to ``Relational Praxiology'':} The theory must ultimately point toward a new philosophy of practice. This entails exploring: at the societal level, how to consciously foster more generative, resilient, and just \emph{relational network configurations}; at the personal level, how to understand the self as a \emph{``relational achievement''} and seek freedom and responsibility within it; at the civilizational level, how to establish a sustainable relationship of \emph{``verificative co-existence''} with non-human agents, including artificial intelligences and ecosystems.

The essence of the world is not a collection of already-existing things but the drama of relations in the making. This dissertation is not the final act of that drama but an attempt to provide a new grammar for its understanding. We invite the reader, with the same generative and relational spirit, to participate in the refinement, critique, and application of this grammar, to collectively face a future woven from complex relations.

\section{Final Words}
\label{sec:final-words}

\emph{Relational Ontology} invites us to view the world anew: not as a collection of separate objects, but as a dynamic, generative network of relations, in which the stability and change of every node (existence) are interwoven with all others. It reminds us that our freedom, responsibility, consciousness, and even suffering are deeply rooted in the texture of the relations we sustain with all things—including each other, and including the technologies we create. \textbf{Cultivating relations capable of responsible co-evolution may be the most fundamental ethical and existential task of our time.}
